SKETCH RECOGNITION AND UNDERSTANDING

### 2009-10-13 trunk/draft/SketchMaster.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

% During the editing process you will not want to view and print the
% whole document. Uncomment the command below so that only the
% chapters that you want to work on are visible.
%\includeonly{ChapterIntroduction}
%\documentstyle[11pt,unhthesis]{report}
\documentclass[11pt,doublespace]{Sketchthesis}

% Extra modules can be added to your document to extend the things
% the Latex can Do. One of these is psfig which allows you to print
% postscript files. You tell Latex to read these packages using the
% \usepackage command:

\usepackage{psfig}
\usepackage{named}
\usepackage{natbib}
\bibliographystyle{plain}
\hyphenation{gno-mon-ly}

\usepackage {graphicx}
\usepackage{subfigure}
\usepackage{longtable}
\usepackage{array}
%\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}

@@

\begin{document}
                                                        % TITLE PAGE
                                                        %======================
\title{Sketch Recognition and Understanding}             % My theses title  using Particle Swarm Optimization Algorithm
\author{Maha Mohammed Nabeel El Meseery}                               % Y name
\prevdegrees{B.S.,Computer  Engineering}       		% Your old degree
\major{Computer Engineering}                            % My current major
\degree{Master of Science}                              % My new degree
\degreemonth{}                                      % When awarded. to be added when known
\degreeyear{}                                      %
\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
%th today or first day of writing  December 26 2005 
%\thesisdate[]                                     % Date of document.
\DOCUMENTtype{Thesis}                                   % or DISSERTATION
\Documenttype{Thesis}                                   % or Dissertation
\documenttype{Thesis}                                   % or dissertation

@@

                                                        % COPYRIGHT PAGE
                                                        %======================
%\copyrightyear{2009}                                    % Delete these
%\makecopyright                                          % if no copyright
                                                    % APPROVAL                                                         %======================
%\thispagestyle{empty}               % Don't do headers or footers.
%\newpage                                                             
%\begin{center}     
%\LARGE  
%\textbf{Sketch Recognition using Particle Swarm Optimization Algorithm}\\
%\normalsize by      \\ 
%\Large  \textbf{ Maha mohamed Nabeel El Meseery}\\
%\vspace*{0.4in}        
%\normalsize
%A Thesis Submitted To the\\
%Faculty of Engineering, Cairo University\\
%in Partial Fulfillment of the\\
%Requirements for the Degree of\\
%MASTER OF SCIENCE
%in\\
%COMPUTER ENGINEERING\\
%
%\end{center}
%          
%    Approved by\\                                           
%Examining Committee:\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Nevin Darwish, Thesis Main Advisor\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Samia Mashali,Thesis Advisor \\
% \makebox[6in]{\hrulefill}\\  
%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
% \makebox[6in]{\hrulefill}\\ 
%Professor  Ashraf Abd El Wahab, Member\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Mohamed Zaki Abd ElMegeed, Member\\
% \makebox[6in]{\hrulefill}\\ 
  % \\   you need...\\
%\makeapproval                                           %

@@

 I thank my family and friends for their moral and support. I specifically thank, dina said, noha mansour and marwah shafee. Thanks, for being always there when I needed you the most. I hope to express my gratitude to all of you for supporting me and praying for me through all these years.
 %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider
%as my sisters. I would like to thank particularly Radwa Aboudina, Fatma Nada, Maha
%Nabil, Haidi Badr, and Marwa Kamal. Thank you all for being always there when I needed
%you most. Thank you for believing in me and supporting me through all these years. I think
%without your support and your prayers, none of this work would be accomplished.
%ii
 \end{acknowledgments}                                   %

                                                        % FOREWORD PAGE
%\begin{foreword}                                        %======================
    											                             % Delete these if
%\end{foreword}                                          % no foreword page.

>>>>

% During the editing process you will not want to view and print the
% whole document. Uncomment the command below so that only the
% chapters that you want to work on are visible.
%\includeonly{ChapterIntroduction}
%\documentstyle[11pt,unhthesis]{report}
\documentclass[11pt,doublespace]{SketchThesis}

% Extra modules can be added to your document to extend the things
% the Latex can Do. One of these is psfig which allows you to print
% postscript files. You tell Latex to read these packages using the
% \usepackage command:
%\usepackage{setspace}
\usepackage{psfig}
\usepackage{named}
\usepackage{natbib}
\bibliographystyle{plain}
\hyphenation{gno-mon-ly}
%\usepackage{setspace}
\usepackage {graphicx}
\usepackage{subfigure}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{array}
\usepackage{algorithmic}
\usepackage{algorithm} 
%\numberwithin{algorithm}{chapter}
%\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}

@@

\begin{document}
                                                        % TITLE PAGE
                                                        %======================
\title{SKETCH RECOGNITION AND UNDERSTANDING}            % My theses title  using Particle Swarm Optimization
% Algorithm   Sketch Recognition and Understanding
\author{Maha Mohamed Nabeel El Meseery Moustafa El Meseery}                               % Y name
\prevdegrees{B.S.,Computer Engineering}       		% Your old degree
\major{COMPUTER ENGINEERING}                            % My current major
\degree{MASTER OF SCIENCE}                              % My new degree
\degreemonth{ }                                      % When awarded. to be added when known
\degreeyear{2009}      
\FacultyUniversity{Faculty of Engineering, Cairo University \\ Giza, Egypt \\  2009}
\superad{}                                %
%\superad{Professor Nevin Darwish\\Professor Samia Mashali\\Assoc. Professor Magda Fayek\\}
%th today or first day of writing  December 26 2005 
%\thesisdate[]                                     % Date of document.
\DOCUMENTtype{ Thesis }                                   % or DISSERTATION
\Documenttype{ Thesis }                                   % or Dissertation
\documenttype{ Thesis }                                   % or dissertation

@@

                                                        % COPYRIGHT PAGE
                                                        %======================
%\copyrightyear{2009}                                    % Delete these
%\makecopyright                                          % if no copyright
  %=======================================Supervisors page ===============================
\newpage    
\thispagestyle{empty}               % Don't do headers or footers.                                                  
 \begin{center}     
 SKETCH RECOGNITION AND UNDERSTANDING\\
\normalsize by      \\ 
  Maha Mohamed Nabeel El Meseery Moustafa El Meseery\\
\vspace*{0.4in}        
\normalsize
A Thesis Submitted To the\\
Faculty of Engineering, Cairo University\\
in Partial Fulfillment of the\\
Requirements for the Degree of\\
MASTER OF SCIENCE
in\\
COMPUTER ENGINEERING\\
\vspace*{0.4in} 
Under the Supervision of 
\end{center}
  \vspace*{0.4in} 
\begin{center}
 		\begin{tabular}{ccc}
		Dr. Nevin Darwish  & Dr. Samia Mashali & Dr. Magda Fayek \\
		 Professor   & Professor & Associate Professor  \\
		 Faculty of Engineering& Electronic Research Institute & 		 Faculty of Engineering \\
		  Cairo University && Cairo University\\
		\end{tabular}
\end{center}
  \vspace*{0.4in}       
  \vfill
\begin{center}
   Faculty of Engineering, Cairo University \\ Giza, Egypt \\ 2009
\end{center}

  %=======================================Supervisors page ===============================
                                                     % APPROVAL                                                         %======================
\newpage       
\thispagestyle{empty}               % Don't do headers or footers.

\begin{center}     
SKETCH RECOGNITION AND UNDERSTANDING\\
\normalsize by      \\ 
Maha Mohamed Nabeel El Meseery Moustafa El Meseery\\
\vspace*{0.2cm}        
A Thesis Submitted To the\\
Faculty of Engineering, Cairo University\\
in Partial Fulfillment of the\\
Requirements for the Degree of\\
MASTER OF SCIENCE
in\\
COMPUTER ENGINEERING\\
\end{center}         
Approved by\\                                           
Examining Committee:\\ \vspace*{0.2cm} 
\begin{tabular*}{0.9\textwidth}{l   }
 %\makebox[5in]{\hrulefill}\\ 
\\ \hline 
Professor Nevin Darwish, Thesis Main Advisor\\ \\  \hline  
 %\makebox[5in]{\hrulefill}\\ 
Professor Samia Mashali,Thesis Advisor \\  \\ \hline   
 %\makebox[6in]{\hrulefill}\\  
Assoc. Professor Magda Fayek,Thesis Advisor\\  \\  \hline  
 %\makebox[5in]{\hrulefill}\\ 
Professor  Ashraf Abd El Wahab, Member\\ \\ \hline  
 %\makebox[6in]{\hrulefill}\\ 
Professor Mohamed Zaki Abd ElMegeed, Member\\  
 %\makebox[6in]{\hrulefill}\\ 
 %\end{spacing}

		\end{tabular*}

%\begin{spacing}{2}
% \makebox[5in]{\hrulefill}\\ 
%Professor Nevin Darwish, Thesis Main Advisor\\
% \makebox[5in]{\hrulefill}\\ 
%Professor Samia Mashali,Thesis Advisor \\
% \makebox[6in]{\hrulefill}\\  
%Assoc. Professor Magda Fayek,Thesis Advisor\\ 
% \makebox[5in]{\hrulefill}\\ 
%Professor  Ashraf Abd El Wahab, Member\\
% \makebox[6in]{\hrulefill}\\ 
%Professor Mohamed Zaki Abd ElMegeed, Member\\
% \makebox[6in]{\hrulefill}\\ 
 %\end{spacing}
 \begin{center}
   Faculty of Engineering, Cairo University \\ Giza, Egypt \\  2009
\end{center}
  % \\   you need...\\
%\makeapproval                                           %

@@

 I thank my family and friends for their moral and support. I specifically thank, Dina Said, Noha Mansour and Marwah Shafee. Thanks, for being always there when I needed you the most. I hope to express my gratitude to all of you for supporting me and praying for me through all these years.
 %Finally, I am very grateful to my dear parents, my family, and my friends whom I consider
%as my sisters. I would like to thank particularly Radwa Aboudina, Fatma Nada, Maha
%Nabil, Haidi Badr, and Marwa Kamal. Thank you all for being always there when I needed
%you most. Thank you for believing in me and supporting me through all these years. I think
%without your support and your prayers, none of this work would be accomplished.
%ii
 \end{acknowledgments}                                   %

                                                        % FOREWORD PAGE

% didn't finish it all Need some addint
The need for new level of human computer interaction has lead researchers to attempt improve the current interfaces. Even though the area of research progressed quickly in speech and handwriting systems, gesture and sketch understanding systems are still starting to get noticed.

Sketch recognition is defined as the process of identifying the symbols the user draws using single or multiple strokes. Users draw strokes using a pen and the system immediately interprets their strokes into objects that can be easily manipulated. This research uses Particle Swarm Optimization algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. Two algorithms are used to divide strokes, the first algorithm \textit{ALgS1} divides strokes to polygons. The second one \textit{AlgS2}, divides strokes to segments of either lines or curves. The research focuses on the effect of PSO segmentations on the sketch recognition process. The final recognition is achieved using a set of geometrical and global shape properties features. A SVM (support vector machine) classifier is used to correctly identify shapes.

Experiments were conducted on three different datasets; Hs-DB a benchmark dataset which contains simple presentation symbols, a LD-DB a set of logic design symbols and EL-DB Electrical symbols. Results show that using \textit{ALgS2} to segment strokes improves segmentation as compared to other algorithms. To be specific \textit{AlgS2} achieves good performance with datasets like EL-DB and Hs-DB however is not as good in datasets characterized by a higher number of curves on every symbolic such as LD-DB.
  %\textit{ALgS1} achieved good performance on EL-DB and Hs-DB but performed poorly on LD-DB due to high number of curves in every symbol
. %The segmentation improvement is tested by final symbol recognition accuracy.

\end{abstractpage}                              % This ends the page.
%\begin{foreword}                                        %======================
    											                             % Delete these if
%\end{foreword}                                          % no foreword page.
